---
title: "Gov 50 Final Project"
author: "Emma Rogers"
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(infer)
```

## Introduction

What factors or characteristics most affect the selling price of used cars? I hypothesize that the distance traveled by a car will have the greatest effect on the selling price and as the distance traveled by a car increases, the cost of this car will decrease. Cars that have been driven more kilometers are worth less than those that have not. This is because the distance a car has traveled indicates how much work and repair the car might need and how much more life the next owner will get out of the car. To examine this relationship, I will be using data scraped from CarDeckho, India’s leading car search website. All variables are measured from the vehicle information included on used car sales listings on CarDeckho in the year 2020. My dependent variable is the selling price of cars, and the independent variables include kilometers driven (continuous), fuel type (binary: Petrol, Diesel) and car age/years since car was first purchased (continuous). The research design for this study is cross-sectional as it is examining the outcome and explanatory variables at one point in time.


```{r, echo = FALSE}
cars <- read_csv("CAR DETAILS FROM CAR DEKHO.csv")
cars <- cars |>
  filter(fuel %in% c("Diesel", "Petrol")) |>
  mutate(age = 2020 - year)
```


```{r, echo = FALSE}
## summarize dependent variable
cars |>
  ggplot(mapping = aes(x = selling_price)) +
  geom_histogram(color = "black", fill = "darkslategray1") +
  labs(x = "Price of Used Car (in dollars)",
       y = "Number of Cars Sold",
       title = "Selling Prices of Used Cars") +
  geom_vline(aes(xintercept = mean(selling_price)), color = "indianred") +
  geom_vline(aes(xintercept = median(selling_price)), color = "blue1") +
  theme_classic() 
```
```{r, echo = FALSE}
## summarize price by fuel type
cars |>
  ggplot(mapping = aes(x = selling_price)) +
  geom_histogram(color = "black", fill = "darkslategray1") +
  labs(x = "Price of Used Car (in dollars)",
       y = "Number of Cars Sold",
       title = "Selling Prices of Used Cars") +
  geom_vline(aes(xintercept = mean(selling_price)), color = "indianred") +
  geom_vline(aes(xintercept = median(selling_price)), color = "blue1") +
  theme_classic() +
  facet_wrap(~ fuel)
```


```{r, echo = FALSE, results= 'hide'}
library(broom)

## Regression: price and km_driven
fit_1 <- lm(selling_price ~ km_driven, data = cars)
fit_1
glance(fit_1)

## Correlation
cor_pr_dist <- cor(cars$selling_price, cars$km_driven)

## Scatterplot 1
cars |>
  ggplot(mapping = aes(x = km_driven, y = selling_price)) +
  geom_point(alpha = 0.5) +
  ylim(0, 10000000) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Distance Traveled by Car (in Kilometers)",
       y = "Price of Car",
       title = "Relationship between distance traveled by car and selling price") +
  theme_bw()

```


```{r, echo = FALSE, results= 'hide'}
## Fit 1 w/out outliers
cars_edit <- cars |>
  filter(km_driven < 5e+05 & selling_price < 7500000)
edit_fit <-
  lm(selling_price ~ km_driven, data = cars_edit)
edit_fit

## scatterplot 1 w/out outliers
cars_edit |>
  ggplot(mapping = aes(x = km_driven, y = selling_price)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Distance Traveled by Car (in Kilometers)",
       y = "Price of Car",
       title = "Relationship between distance traveled and selling price (w/out outliers)") +
  theme_bw()
```
There is a negative correlation between distance traveled and the selling price at a value of -0.19.

```{r, echo = FALSE, results= 'hide'}
## Multiple Regression: km_driven + fuel
mult_fit1 <- lm(selling_price ~ km_driven + fuel, data = cars)
mult_fit1
glance(mult_fit1)
```

```{r, echo = FALSE, results= 'hide'}
## Mult Regression: km_driven + age
mult_fit2 <- lm(selling_price ~ km_driven + age, data = cars)
mult_fit2
```

```{r, echo = FALSE, results= 'hide'}
## Multiple Regression: km_driven + fuel + age
mult_fit4 <- lm(selling_price ~ km_driven + fuel + age, data = cars)
mult_fit4
glance(mult_fit4)
```

```{r, echo = FALSE, results= 'hide'}
## Regression w/ age
fit_age <- lm(selling_price ~ age, data = cars)
fit_age
glance(fit_age)

cor_pr_age <- cor(cars$selling_price, cars$age)

```

```{r, echo = FALSE, results = 'hide'}
mult_fit3 <- lm(selling_price ~ fuel + age, data = cars)
mult_fit3
```


```{r, echo = FALSE, results= 'hide'}
## Scatterplot for age and price
cars |>
  ggplot(mapping = aes(x = age, y = selling_price)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Years since car was purchased",
       y = "Price of Car",
       title = "Relationship between car age and selling price") +
  theme_bw()
```
There is a negative correlation between age and selling price at a value of -0.41.

```{r, echo= FALSE, results = 'hide'}
## correlation between age and dist
cor_dist_age <- cor(cars$km_driven, cars$age)

## scatterplot between age and dist
cars |>
  ggplot(mapping = aes(x = km_driven, y = age)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Distance Traveled by Car (in Km)",
       y = "Years since car was purchased",
       title = "Relationship between distance traveled and car age") +
  theme_bw()
```
There is a positive correlation of 0.42 between two of our independent variables: age and km_driven.

```{r, echo = FALSE}
## Regressions Table
modelsummary::modelsummary(
  statistic = c("s.e. = {std.error}",
               "p = {p.value}"),
  list(fit_1, mult_fit1, mult_fit2, mult_fit3, mult_fit4),
  gof_map = c("nobs", "r.squared", "adj.r.squared")) 
```

## Results

```{r, echo = FALSE}
cars |>
  ggplot(mapping = aes(x = age, y = selling_price)) +
  geom_point(mapping = aes(color = km_driven), alpha = 0.5) +
  scale_color_distiller(palette = "Spectral") +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Years since car was purchased",
       y = "Price of Car",
       title = "Main regression") +
  theme_bw() +
  facet_wrap(~ fuel)
```

```{r}
## Main Regression of interest
modelsummary::modelsummary(
  statistic = c("s.e. = {std.error}",
               "p = {p.value}"),
  list(mult_fit4),
  gof_map = c("nobs", "r.squared", "adj.r.squared"))
```




