---
title: "Gov 50 Final Project"
author: "Emma Rogers"
description: "Prices of Used Cars"
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(infer)
```

## Introduction

What factors or characteristics most affect the selling price of used cars? I hypothesize that the distance traveled by a car will have the greatest effect on the selling price and as the distance traveled by a car increases, the cost of this car will decrease. Cars that have been driven more kilometers are worth less than those that have not. This is because the distance a car has traveled indicates how much work and repair the car might need and how much more life the next owner will get out of the car. To examine this relationship, I will be using data scraped from CarDeckho, India’s leading car search website. All variables are measured from the vehicle information included on used car sales listings on CarDeckho in the year 2020. My dependent variable is the selling price of cars, and the independent variables include kilometers driven (continuous), fuel type (binary: Petrol, Diesel) and car age/years since car was first purchased (continuous). The research design for this study is cross-sectional as it is examining the outcome and explanatory variables at one point in time.


```{r, echo = FALSE}
## Loading Data
cars <- read_csv("CAR DETAILS FROM CAR DEKHO.csv")
cars <- cars |>
  filter(fuel %in% c("Diesel", "Petrol")) |>
  mutate(age = 2020 - year)
```


```{r, echo = FALSE}
## summarize dependent variable
cars |>
  ggplot(mapping = aes(x = selling_price)) +
  geom_histogram(color = "black", fill = "darkslategray1") +
  labs(x = "Price of Used Car (in dollars)",
       y = "Number of Cars Sold",
       title = "Selling Prices of Used Cars",
       tag = "Figure 1:") +
  geom_vline(aes(xintercept = mean(selling_price)), color = "indianred") +
  geom_vline(aes(xintercept = median(selling_price)), color = "blue1") +
  theme_classic() 
```
```{r, echo = FALSE}
## summarize price by fuel type
cars |>
  ggplot(mapping = aes(x = selling_price)) +
  geom_histogram(color = "black", fill = "darkslategray1") +
  labs(x = "Price of Used Car (in dollars)",
       y = "Number of Cars Sold",
       title = "Selling Prices of Used Cars",
       tag = "Figure 2:") +
  geom_vline(aes(xintercept = mean(selling_price)), color = "indianred") +
  geom_vline(aes(xintercept = median(selling_price)), color = "blue1") +
  theme_classic() +
  facet_wrap(~ fuel)
```


```{r, echo = FALSE, results= 'hide'}
library(broom)

## Correlation between price and distance
cor_pr_dist <- cor(cars$selling_price, cars$km_driven)

## Scatterplot 1
cars |>
  ggplot(mapping = aes(x = km_driven, y = selling_price)) +
  geom_point(alpha = 0.5) +
  ylim(0, 10000000) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Distance Traveled by Car (in Kilometers)",
       y = "Price of Car",
       title = "Correlation: Distance Traveled By Car vs Selling Price",
       tag = "Figure 3:") +
  theme_bw()

```


```{r, echo = FALSE, results= 'hide'}
## Figure 3 w/out outliers
cars_edit <- cars |>
  filter(km_driven < 5e+05 & selling_price < 7500000)
edit_fit <-
  lm(selling_price ~ km_driven, data = cars_edit)
edit_fit

## scatterplot 1 w/out outliers
cars_edit |>
  ggplot(mapping = aes(x = km_driven, y = selling_price)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Distance Traveled by Car (in Kilometers)",
       y = "Price of Car",
       title = "Correlation: Distance Traveled vs Selling Price (w/out outliers)",
       tag = "Figure 4:") +
  theme_bw()
```
There is a negative correlation between distance traveled and the selling price at a value of -0.19.


```{r, echo = FALSE, results= 'hide'}
## correlation w/ age
cor_pr_age <- cor(cars$selling_price, cars$age)

## Scatterplot for age and price
cars |>
  ggplot(mapping = aes(x = age, y = selling_price)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Years since car was purchased",
       y = "Price of Car",
       title = "Correlation: Car Age vs Selling Price",
       tag = "Figure 5:") +
  theme_bw()
```
There is a negative correlation between age and selling price at a value of -0.41.

```{r, echo= FALSE, results = 'hide'}
## correlation between age and dist
cor_dist_age <- cor(cars$km_driven, cars$age)

## scatterplot between age and dist
cars |>
  ggplot(mapping = aes(x = km_driven, y = age)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Distance Traveled by Car (in Km)",
       y = "Years since car was purchased",
       title = "Correlation: Distance Traveled By Car vs Car Age",
       tag = "Figure 6:") +
  theme_bw()
```
There is a positive correlation of 0.42 between two of our independent variables: age and km_driven.

```{r, echo = FALSE}
## Regressions Table
var_labels <- c(
  "(Intercept)" = "Intercept",
  "km_driven" = "Km Driven",
  "fuelPetrol" = "Petrol Fuel vs Diesel",
  "age" = "Age"
)

modelsummary::modelsummary(
  statistic = c("s.e. = {std.error}",
               "p = {p.value}"),
  list(
    "Model 1" = lm(selling_price ~ km_driven, data = cars),
    "Model 2" = lm(selling_price ~ km_driven + age, data = cars), 
    "Model 3" = lm(selling_price ~ age, data = cars),
    "Model 4" = lm(selling_price ~ fuel + age, data = cars),
    "Model 5" = lm(selling_price ~ km_driven + fuel + age, data = cars)),
  coef_map = var_labels,
  title = "Regression Table",
  gof_map = c("nobs", "r.squared", "adj.r.squared"))

```

## Results

```{r, echo = FALSE}
cars |>
  ggplot(mapping = aes(x = age, y = selling_price)) +
  geom_point(aes(color = fuel), alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Years since car was purchased",
       y = "Price of Car",
       title = "Main Regression of Interest (Model 4)",
       tag = "Figure 7:",
       color = "Fuel") +
  theme_bw()
```

```{r, echo = FALSE}
## Main Regression of interest
modelsummary::modelsummary(
  statistic = c("s.e. = {std.error}",
               "p = {p.value}"),
  list(
    "Model 4" = lm(selling_price ~ fuel + age, data = cars)),
  title = "Main Regression of Interest",
  coef_map = var_labels,
  gof_map = c("nobs", "r.squared", "adj.r.squared"))
```


Although the kilometers driven by a car does have a negative correlation with the selling price, age of the car has the greatest effect on the outcome variable. While keeping fuel constant, a 1 year increase in age is associated with a $53,247.31 decrease in the selling price of a car. With a test level of 0.05, the estimated coefficient of age is statistically significant as the p-value is approximately zero. Though there is a significant correlation of 0.42 between age and the price of a used car, this relationship cannot be interpreted causally. This is due to the multiple possible confounders that are not measured in this dataset that could be affecting the selling price of a car. Possible confounders include the exterior condition of the car, special features and technology, and number of seats. Without accounting for these other variables and their impacts on the outcome variable, we can not conclude that the coefficient of car age represents a causal effect.

## Conclusion

Through analyzing this data, we found that distance traveled by a car does have a negative correlation with its selling price. However, after running regressions accounting for multiple independent variables, the distance traveled had a lesser effect on selling price than I had expected.

First, figure 3 displays the relationship between the independent variable km_driven and the outcome variable selling price. Figure 4 is a replication of figure 3 with some extreme outliers filtered out to see this relationship more clearly. This scatterplot displays a negative correlation of -0.19 between distance traveled and selling price indicating that a 1 Km increase in distance traveled is associated with a decrease in selling price of $2.37. The correlation coefficient for km_driven has a p-value of approximately 0 which is statistically significant at a test level of 0.05. However, the R squared value for this simple regression was 0.036, a very low value that suggests that this model does not fit the data well. After exploring this initial relationship, I completed a number of regressions with different combinations of variables to compare the effects each variable had on the outcome. In Model 2, I accounted for both km driven and age of a car and the R squared value rose to 0.171. While keeping km driven constant, an increase in 1 year of age is associated with a $55,887.30 decrease in price at a p-value of 0. With the addition of such a significant independent variable, the effect of km driven on selling price holding age constant decreased to -0.259 at a p-value of 0.174. By including more information in the regression, model 2 displays how age is a lot more important than distance. Figure 5 demonstrates the stronger correlation between age and selling price of -0.41. Then, I ran a simple regression with age and selling price and a 1 year increase in age is associated with a decrease in selling price of $57094.54. This model has the same R squared value as model 2 of 0.171 which means that km driven does not add or help this model. By calculating and visualizing the relationship between distance traveled and car age in Figure 6, I discovered there is a positive correlation of 0.42 and therefore some colinearity affecting these regressions. This makes logical sense because the longer an owner has a car, the more likely they are to have driven more kilometers. This might be impacting our interpretation of the effects of each individual variable because we cannot be sure that our coefficient represents a case when the other independent variable is constant. Because of the redundancy of this collinearity, we can remove km driven without strongly affecting the fit of the model. Next, I ran a multiple regression with fuel and age as independent variables and the R squared value increased to 0.224 which displays that fuel type has a significant effect on the selling price (Model 4). While holding age constant, the selling price associated with Petrol fuel is $270,434.90 less than that of Diesel fuel. While holding fuel constant, a 1 year increase in age is associated with a $53,247.31 decrease in selling price. Finally, I tried a regression with all three independent variables, and the performance only increased by 0.014 (Model 5). This confirmed my decision to use Model 4 and leave out km driven because according to Ocaam’s Razor it is better to use the simpler model and only use the independent variables that improve the model’s performance. 

In this analysis, a threat to inference is omitted variable bias because we are leaving out a variable (km driven) which is statistically significant on its own. Therefore, our results might not be accurate because we are not accounting for a variable that does have some effect on the outcome. However, because of the collinearity between age and km driven, it is better to leave out this variable as collinearity is a larger threat to inference than omitting a variable. In future work, this analysis could be improved by collecting more information on the car. As I stated above, there are many possible confounders that are not measured in this data set such as the exterior appearance of the car that would most likely have strong impacts on the selling price of a car. With more information about the car, we can decrease the amount of bias and come closer to identifying a causal effect.
